Microlensing as a possible probe of event-horizon structure in quasars

### 2016-11-23 paper1/msold2.tex
template for response2

Gravitational microlensing of quasars, reviewed in
Section~\ref{sec:microlensing} below, is a phenomenon seen in quasars
that have been multiply imaged by galaxies.  Individual images in a
multiple-image system can undergo sharp and seemingly random
brightness changes.  This can occur, even for constant intrinic
luminosity, as a consequence of two things: the very small size of the central engine,
and granularity of the mass distribution of a lensing galaxy due to
stars.  The latter means that the local magnification is not a smooth
function of source position. It contains a complicated network of
singular curves, known as caustics.
Figure~\ref{fig:magnification_map} shows part of a magnification map
with a few caustics.  The lensed brightness would be given by placing
the source on such a magnification map and integrating the surface
brightness weighted by the magnification.  Most astrophysical sources
straddle several caustics, and hence, their net brightness varies
smoothly with location.  The central engine of quasars, however, is
smaller than the typical spacing between caustics.  As a result, the
lensed brightness undergoes sudden changes as a quasar crosses a
caustic.  Quasar microlensing is mostly observed in the optical
continuum \citep[e.g.,][]{2012A&A...544A..62S} but is also known in
X-rays \citep[e.g.,][]{2016arXiv160207601N} and has been argued for at
cm wavelengths \citep{2000A&A...358..793K}.  The effect supplies an
upper limit on the size of the central engine
\citep[e.g.,][]{2015ApJ...814L..26M} and can be used to study the mass
distribution and kinematics of stars in the lensing galaxy as well
\citep[e.g.,][]{2012ApJ...744..111P}.  Microlensing caustics have an
additional remarkable feature: though they can be very complicated,
they have some universal properties well-known from catastrophe
theory.  In particular, very close to the simplest caustics (known as
folds), the magnification is approximately constant on one side and
$\propto1/\sqrt p$ where $p$ is the transverse distance of the source
from the caustic.  This property will be exploited later.

>>>>

Gravitational microlensing of quasars, reviewed in
Section~\ref{sec:microlensing} below, is a phenomenon seen in quasars
that have been multiply imaged by galaxies.  Individual images in a
multiple-image system can undergo sharp and seemingly random
brightness changes.  This can occur, even for constant intrinic
luminosity, as a consequence of two things: the very small size of the
central engine, and granularity of the mass distribution of a lensing
galaxy due to stars.  The latter means that the local magnification is
not a smooth function of source position. It contains a complicated
network of singular curves, known as caustics.
Figure~\ref{fig:magnification_map} shows part of a magnification map
with a few caustics.  The lensed brightness would be given by placing
the source on such a magnification map and integrating the surface
brightness weighted by the magnification.  Most astrophysical sources
straddle several caustics, and hence, their net brightness varies
smoothly with location.  The central engine of quasars, however, is
smaller than the typical spacing between caustics.  As a result, the
lensed brightness undergoes sudden changes as a quasar crosses a
caustic.  Quasar microlensing is mostly observed in the optical
continuum \citep[e.g.,][]{2012A&A...544A..62S} but is also known in
X-rays \citep[e.g.,][]{2016arXiv160207601N}.  In principle, it exists
at all wavelengths \citep[and has been argued for at cm wavelengths
  by][]{2000A&A...358..793K} but will be washed out if the light is
from larger regions than the central engine of the quasar.  The effect
thus supplies an upper limit on the size of the central engine
\citep[e.g.,][]{2015ApJ...814L..26M} and can be used to study the mass
distribution and kinematics of stars in the lensing galaxy as well
\citep[e.g.,][]{2012ApJ...744..111P}.  Microlensing caustics have an
additional remarkable feature: though they can be very complicated,
they have some universal properties well-known from catastrophe
theory.  In particular, very close to the simplest caustics (known as
folds), the magnification is approximately constant on one side and
$\propto1/\sqrt p$ where $p$ is the transverse distance of the source
from the caustic.  This property will be exploited later.
